Making multiple-locus association mapping on a genome-wide scale routine

### 2017-11-08 AM+draft.tex
09Nov17 morning - about to work on methods

We compared the performance of Eagle against five other multiple locus association mapping methods: BigRR, 
glmnet, FaST-LASSO, MMLM, and r2VIM.  These five methods were selected because of their demonstrated value 
for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
features is given in Table X.

Of the five multiple locus methods considered, BigRR, LMM-Lasso, and glmnet,  implement regularization methods for parameter
estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
genomic prediction (refs) but also association mapping (refs). BigRR is an R package that implements generalized ridge regression
which is an improvement on standard ridge regression. In standard ridge regression, coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.

>>>>

We compared the computational and statistical performance of Eagle against five other multiple locus applications for 
association mapping  These were BigRR, LMM-Lasso, glmnet, MLMM  and, r2VIM. 
BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. Regularisation methods have dominated 
the development of multiple locus approaches.  They give opportunity for the effect of all the snp to be modelled simultaneously. 
By introducing a penalty function that balances bias and variance, regularisation methods can 
handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
MLMM is similar in philosophy to Eagle in that it treats association mapping as a model building problem. Here, a linear mixed 
model is built through a forward-backward selected strategy.  
R2VIM is different to the others in that it implements 
random forests for association mapping. 
Random forests is a machine learning approach that forms an ensemble of decision trees for regression and 
classification. 
These five methods were selected because of their demonstrated value 
for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
features is given in Table X. Each method is now described in a little more detail.

Regression based regularization methods 
are well suited to solving problems where the number of predictors (p) is much greater than the number of samples (as in
genome-wide association studies).  Unsurprisingly,  regularization methods have dominated the development of 
multiple locus assciation mapping approaches.  Regularisation-based 
approaches model the effect of all the snp on a trait simultaneously. 
Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
bias against reduced variance. The fourth application implements a forward-backward selection approach for building a linear 
mixed model. The fifth application uses random forests. 
These five methods were selected because of their demonstrated value 
for analysing genetic data, they reflect a range of different statistical methodologies for association mapping, and they are available as 
either stand-alone computer programs or as  packages. A summary of how the computer programs/packages differ in their 
features is given in Table X. Each method is now described briefly.

BigRR is an R package that implements generalized ridge regression. Generalized ridge regression is 
which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.

BigRR, LMM-Lasso, and glmnet all implement regularization methods for parameter
estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
genomic prediction (refs) but also association mapping (refs). 
BigRR implements generalized ridge regression, LMM-Lasso implements a two-step lasso approach, and glmnet makes use of 
elastic net for parameter estimation.

BigRR is an R package that implements generalized ridge regression
which is an improvement on ridge regression. In ridge regression, coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.
